|

### 2024-10-16 posts/BarthPapageorgeThom/BPT.qmd
Added gene-wealth gradient with education control in BPT.html

@@

$$
Y_{i}=\bfmu'\bfx_{i}+\beta_{j}SNP_{ij}+\epsilon_{ij}, \quad j=1,\dots,J.
$$

@@

>>>>

    css: ../seiro.css

@@

$$
Y_{i}
=
\beta_{j}SNP_{ij}+\bfmu'_{j}\bfx_{i}+\epsilon_{ij}, \quad j=1,\dots,J.
$$

$\cov[SNP_{ij}, \epsilon_{ij}]> 0$: Parents see traits which may be correlated with $SNP_{ij}$ and decide on investments. Environment is endogenous to the trait in SNP regressions.

@@

Gene-wealth gradient $\gamma$ with education control

$$
\begin{aligned}
W_{i}
&=
\bftheta'\bfx_{i}+\gamma PGS_{i}+\eta Y_{i}+e_{i},\\
&=
\bftheta'\bfx_{i}+\gamma PGS_{i}+\frac{\eta}{J}\left\{PGS_{i}+\sum_{j=1}^{J}\left(\bfmu'_{j}\bfx_{i}+\tilde{\epsilon}_{ij}\right)\right\}+e_{i},\\
&=
\bftheta'\bfx_{i}+\left(\gamma+\frac{\eta}{J}\right)PGS_{i}+\frac{\eta}{J}\sum_{j=1}^{J}\left(\bfmu'_{j}\bfx_{i}+\tilde{\epsilon}_{ij}\right)+e_{i},\\
&\simeq
\left(\bftheta'+\eta\bar{\bfmu}'\right)\bfx_{i}+\gamma PGS_{i}+\eta\bar{\tilde{\epsilon}}_{i}+e_{i}.
\end{aligned}
$$
where
$$
JY_{i}
=
PGS_{i}+\sum_{j=1}^{J}\left(\bfmu'_{j}\bfx_{i}+\tilde{\epsilon}_{ij}\right).
$$

Because $\cov[SNP_{ij}, \epsilon_{ij}]> 0$, in general, we have $\cov[PGS_{i}, \bar{\tilde{\epsilon}}_{i}]> 0$.

So, in general, $\plim \hat{\gamma}-\gamma>0$.

This is problematic: the paper tries to show persistence of positive $\hat{\gamma}$

Authors may have acknowledged it: Kept on cautioning that

* Estimated relationship is not causal
* Diverse mechanisms give rise to estimated relationship

A partial remedy: Use $\tilde{Y}_{i}=\tilde{\beta}_{j}SNP_{ij}+\bfmu'_{j}\bfx_{i}$ in place of $Y_{i}$ on RHS.

This *partly* takes away the correlation between $GPS_{i}$ and the error term

Because $\cov[SNP_{ij}, \epsilon_{ij}]> 0$, $\tilde{\beta}_{j}$ is overestimated, so deviation of $\hat{\tilde{Y}}_{i}$ from real $\tilde{Y}_{i}$ is larger for individual with large $SNP_{ij}$. Although partly removed, there still remain a positive correlation between PGS and the error term.
